---
title: "p8105_hw3_hq2163_hanbo"
author: "Hanbo Qiu"
date: "October 7, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
  fig.width = 6,
  fig.asp = .8,
  out.width = "90%"
)

library(tidyverse)
library(ggridges)
library(patchwork)
library(p8105.datasets)

theme_set(theme_bw() + theme(legend.position = "bottom"))
```

##Problem 1

####Create an overall_health dataset and clean it

```{r}
data(brfss_smart2010)

overall_health = janitor::clean_names(brfss_smart2010) %>% 
  select(year, location_abbr = locationabbr, location_desc = locationdesc, everything()) %>%
  filter(topic == "Overall Health") %>% 
  mutate(response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"), ordered = TRUE))
```

We first load the data. Then we format the variables to use appropriate names, focus on the “Overall Health” topic, and organize responses as a factor taking levels ordered from “Excellent” to “Poor”.

####Which states were observed at 7 locations in 2002?

```{r}
#Make a table with distinct locations in 2002.
filter(overall_health, year == 2002) %>% 
  distinct(location_abbr, location_desc) %>% 
  count(location_abbr) %>% 
  filter(n == 7) %>% 
  knitr::kable(col.names = c("State", "No. of locations"))
```

The first column shows the states observed at 7 locations in 2002.

####Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

```{r}
distinct(overall_health, year, location_abbr, location_desc) %>% 
  ggplot(aes(x = year, color = location_abbr)) +
  geom_freqpoly(binwidth = 1) +
  scale_x_continuous(breaks = 2002:2010, limits = c(2002,2010)) +
  labs(
    title = "The number of locations in each state from 2002 to 2010",
    x = "Year",
    y = "No. of locations",
    caption = "Data from the brfss_smart2010"
  ) +
  theme(legend.position = "none")
```

The number of observations in State jumps to above 40 in 2007 and 2010. The number of observations all the other states kept under 20 for the rest of the years.

####Showing the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State in 2002, 2006, and 2010.

```{r}
overall_health %>% 
  filter(year %in% c(2002, 2006, 2010), 
         response == "Excellent", 
         location_abbr == "NY") %>% 
  group_by(year) %>% 
  summarise(mean = mean(data_value, na.rm = TRUE),
            sd = sd(data_value, na.rm = TRUE)) %>% 
  knitr::kable(col.names = c("Year", "Mean of excellent proportion(%)", "SD of excellent proportion(%)"),
               digits = 2)
```

####For each year and state, compute the average proportion and make a five-panel plot.

```{r}
select(overall_health, year, location_abbr, response, data_value) %>% 
  group_by(response, year, location_abbr) %>% 
  summarise(mean = mean(data_value, na.rm = TRUE)) %>% 
  ungroup %>% 
  mutate(year = as.factor(year)) %>% 
  ggplot(aes(x = year, y = mean)) +
  geom_dotplot(binaxis="y", stackdir = "center", stackratio = 2) +
  facet_grid(response ~., scales = "free_y") +
  labs(
    title = "Response distribution of state-level averages from 2002-2010",
    x = "Year",
    y = "The average proportion of response",
    caption = "Data from the brfss_smart2010"
  )
```

##Problem 2

####Load the data from the p8105.datasets package:

```{r}
data(instacart)
```

####Write a short description of the dataset:

Instacart is an online grocery service that allows you to shop online from local stores. The dataset contains `r nrow(instacart)` observations and `r ncol(instacart)` variables of `r distinct(instacart, user_id) %>% nrow()` users. Each row in the dataset shows the information of a specific product in one order. Some key variables are user_id and order_id, by which we can identify each user and order. Product_id/product_name, aisle_id/aisle, department_id/department are key variables for product. The most important variables are reordered, order_number, order_dow, order_hour_of_day, days_since_prior_order, they can help us understand the customers' purchasing and consumption behaviour. 

####How many aisles are there, and which aisles are the most items ordered from?

There are `r distinct(instacart, aisle_id) %>% nrow()` aisles. 

```{r}
item_aisle = count(instacart, aisle_id, sort = T)
```

No. `r item_aisle[1,1]` is the most items ordered aisle.

####Make a plot that shows the number of items ordered in each aisle.



####Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r}
filter(instacart, aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>% 
  summarise(n = n()) %>% 
  filter(min_rank(desc(n)) == 1) %>% 
  select(aisle, product_name) %>% 
  arrange(aisle) %>% 
  knitr::kable(col.names = c("Aisle", "Most popular product"))
```

####Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}
filter(instacart, product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean = round(mean(order_hour_of_day), 0)) %>% 
  mutate(mean = as.character(mean),
         mean = paste(mean, ":00", sep = "")) %>% 
  spread(key = order_dow, value = mean ) %>% 
  knitr::kable(caption = "The mean hour of the day",
               col.names = c("Product", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))
```

The above table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. As the data does not specify the meaning of values in order_dow variable, I presume that the values of 0~6 represent Sunday to Saturday.

##Problem 3

**Load the data**

```{r}
data(ny_noaa)
```

**Write a short description of the dataset:**

The ny_noaa data provides weather records in NY state during 1981-2010. It contains `r nrow(ny_noaa)` observations and `r ncol(ny_noaa)` variables, where each row records the weather records of a certain day observed in a certain sation. The key variables include `prcp`, `snow`, `snwd`, `tmax`, and `tmin`, which represent precipitation, snowfall, snow depth, maximum temperature, and minimun temperature, respectively. `prcp`, `snow` and `snwd` are all `integer` class variables, while `tmax` and `tmin` are `character` variables.

The proportion of missing values in some variables are `r signif(sum(is.na(ny_noaa$prcp)) / nrow(ny_noaa), 2)` for `prcp`, `r signif(sum(is.na(ny_noaa$snow)) / nrow(ny_noaa), 2)` for `snow`, `r signif(sum(is.na(ny_noaa$snwd)) / nrow(ny_noaa), 2)` for `snwd`, `r signif(sum(is.na(ny_noaa$tmax)) / nrow(ny_noaa), 2)` for `tmax`, and `r signif(sum(is.na(ny_noaa$tmin)) / nrow(ny_noaa), 2)` for `tmin`. Each weather station collect only a subset of these variables, and therefore the resulting dataset contains extensive missing data which might influence the result.

**Do some data cleaning:**

```{r}
ny_noaa_ymd = separate(ny_noaa, date, into = c("year", "month", "day"), sep = "-") %>% 
    mutate(prcp = prcp / 10, tmin = as.numeric(tmin) / 10, tmax = as.numeric(tmax) / 10)
```

We separate `date` to three variables (`year`, `month`, `day`), and convert observations for `prcp`, `tmax` and `tmin` to reasonable units.

**For snowfall, what are the most commonly observed values? Why?**

```{r}
snowfall_daycount = count(ny_noaa, snow, sort = T)
head(snowfall_daycount, 3)
```

From the above table, we find the most observed values for snowfall is `r snowfall_daycount[1,1]` mm. The reason is that there is no snow fall in most days in NY.

**Make a two-panel plot showing the average max temperature in January and in July in each station across years:**

```{r}
filter(ny_noaa_ymd, month %in% c("01","07")) %>%
  mutate(month = recode(month, "01" = "Jan", "07" = "Jul")) %>% 
  group_by(id, year, month) %>% 
  summarise(mean_tmax = mean(tmax, na.rm = T)) %>%
  ggplot(aes(x = year, y = mean_tmax)) +
  geom_violin(color = "black", fill = "blue", alpha = 0.4) +
  stat_summary(fun.y = median, geom = "point", color = "black", size = 1.5) +
  scale_x_discrete(breaks = c("1981", "1986", "1991", "1996", "2001", "2006", "2010")) +
  facet_grid(month ~., scales = "free_y") +
  labs(
    title = "Average maximum temperature in Jan and in Jul in each station",
    x = "Year",
    y = "Average maximum daily temperature (C)",
    caption = "Data from the NY NOAA"
  )
```

**Make a two-panel plot (i) `tmax` vs `tmin` for the full dataset; and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year:**

```{r}
tmax_tmin = ggplot(ny_noaa_ymd, aes(x = tmin, y = tmax)) + 
  geom_hex(bins = 40) +
  labs(
    title = "Maximum temperature vs minimum temperature",
    x = "Minimum temperature (C)",
    y = "Maximum temperature (C)"
  ) +
  coord_fixed() +
  theme(legend.position = "right")

snow_fall = filter(ny_noaa_ymd, snow > 0, snow <100) %>% 
  ggplot(aes(x = year, y = snow)) +
  geom_boxplot() +
  scale_x_discrete(breaks = c("1981", "1986", "1991", "1996", "2001", "2006", "2010")) +
  labs(
    title = "The distribution of snowfall in each year",
    x = "Year",
    y = "Snow fall (mm)",
    caption = "Data from the NY NOAA"
  )  

tmax_tmin + snow_fall + plot_layout(ncol = 1, heights = c(3, 1))
```