---
title: "p8105_hw3_hq2163_hanbo"
author: "Hanbo Qiu"
date: "October 7, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

#Load the new package. 
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```

##Problem 1

####Create an overall_health dataset and clean it

```{r}
data(brfss_smart2010)

overall_health = janitor::clean_names(brfss_smart2010) %>% 
  select(year, location_abbr = locationabbr, location_desc = locationdesc, everything()) %>%
  filter(topic == "Overall Health") %>% 
  mutate(response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"), ordered = TRUE))
```

We first load the data. Then we format the variables to use appropriate names, focus on the “Overall Health” topic, and organize responses as a factor taking levels ordered from “Excellent” to “Poor”.

####Which states were observed at 7 locations in 2002?

```{r}
#Make a table with distinct locations in 2002.
filter(overall_health, year == 2002) %>% 
  distinct(location_abbr, location_desc) %>% 
  count(location_abbr) %>% 
  filter(n == 7) %>% 
  knitr::kable(col.names = c("State", "No. of locations"))
```

The first column shows the states observed at 7 locations in 2002.

####Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.

```{r}
distinct(overall_health, year, location_abbr, location_desc) %>% 
  ggplot(aes(x = year, color = location_abbr)) +
  geom_freqpoly(binwidth = 1) +
  scale_x_continuous(breaks = 2002:2010, limits = c(2002,2010)) +
  labs(
    title = "The number of locations in each state from 2002 to 2010",
    x = "Year",
    y = "No. of locations",
    caption = "Data from the brfss_smart2010"
  ) +
  theme(legend.position = "none")
```

The number of observations in State jumps to above 40 in 2007 and 2010. The number of observations all the other states kept under 20 for the rest of the years.

####Showing the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State in 2002, 2006, and 2010.

```{r}
overall_health %>% 
  filter(year %in% c(2002, 2006, 2010), 
         response == "Excellent", 
         location_abbr == "NY") %>% 
  group_by(year) %>% 
  summarise(mean = mean(data_value, na.rm = TRUE),
            sd = sd(data_value, na.rm = TRUE)) %>% 
  knitr::kable(col.names = c("Year", "Mean of excellent proportion(%)", "SD of excellent proportion(%)"),
               digits = 2)
```

####For each year and state, compute the average proportion and make a five-panel plot.

```{r}
select(overall_health, year, location_abbr, response, data_value) %>% 
  group_by(response, year, location_abbr) %>% 
  summarise(mean = mean(data_value, na.rm = TRUE)) %>% 
  ungroup %>% 
  mutate(year = as.factor(year)) %>% 
  ggplot(aes(x = year, y = mean)) +
  geom_dotplot(binaxis="y", stackdir = "center", stackratio = 2) +
  facet_grid(response ~., scales = "free_y") +
  labs(
    title = "Response distribution of state-level averages from 2002-2010",
    x = "Year",
    y = "The average proportion of response",
    caption = "Data from the brfss_smart2010"
  )
```

##Problem 2

####Load the data from the p8105.datasets package:

```{r}
data(instacart)
```

####Write a short description of the dataset:

Instacart is an online grocery service that allows you to shop online from local stores. The dataset contains `r nrow(instacart)` observations and `r ncol(instacart)` variables of `r distinct(instacart, user_id) %>% nrow()` users. Each row in the dataset shows the information of a specific product in one order. Some key variables are user_id and order_id, by which we can identify each user and order. Product_id/product_name, aisle_id/aisle, department_id/department are key variables for product. The most important variables are reordered, order_number, order_dow, order_hour_of_day, days_since_prior_order, they can help us understand the customers' purchasing and consumption behaviour. 

####How many aisles are there, and which aisles are the most items ordered from?

There are `r distinct(instacart, aisle_id) %>% nrow()` aisles. 

```{r}
item_aisle = count(instacart, aisle_id, sort = T)
```

No. `r item_aisle[1,1]` is the most items ordered aisle.

####Make a plot that shows the number of items ordered in each aisle.



####Make a table showing the most popular item aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”.

```{r}
filter(instacart, aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle, product_name) %>% 
  summarise(n = n()) %>% 
  filter(min_rank(desc(n)) == 1) %>% 
  select(aisle, product_name) %>% 
  arrange(aisle) %>% 
  knitr::kable(col.names = c("Aisle", "Most popular product"))
```

####Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}
filter(instacart, product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarise(mean = round(mean(order_hour_of_day), 0)) %>% 
  mutate(mean = as.character(mean),
         mean = paste(mean, ":00", sep = "")) %>% 
  spread(key = order_dow, value = mean ) %>% 
  knitr::kable(caption = "The mean hour of the day",
               col.names = c("Product", "Sun", "Mon", "Tue", "Wed", "Thu", "Fri", "Sat"))
```

The above table shows the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. As the data does not specify the meaning of values in order_dow variable, I presume that the values of 0~6 represent Sunday to Saturday.


##Problem 3

```{r}
# load the data from the p8105.datasets package and summarize the data
data(ny_noaa)

summary(ny_noaa)
```

A short description of the dataset: ny_noaa provides some weather data in NY city, including the GHCN -Daily database of summary statistics from weather stations. The dataset contains 2595176 observations and 7 variables, where each row in the dataset is daily record of weather during 1981-2010.
Key variables includes 1."date" is Date of observation 2."prcp" which represents Precipitation (tenths of mm) recorded by the weather station  on a specific day 3."snow" which represents Snowfall (mm) recorded by the weather station on a specific day 4."snwd": which represents Snow depth (mm) recorded by the weather station on a specific day 5."tmax" which represents the maximum temperature (tenths of degrees C) recorded by the weather station on a specific day 6."tmin" which represents the Minimum temperature (tenths of degrees C) recorded by the weather station on a specific day.
There are missing data in variables: "prcp"(145838 NA), "snow"(381221 NA), "snwd"(591786 NA) and ."tmax","tmin". Each weather station collect only a subset of these variables, and therefore the resulting dataset contains extensive missing data which might influence the result.

Do some data cleaning and create separate variables for year, month, and day.

Convert observations for temperature, precipitation, and snowfall.
```{r}
ny_noaa_ymd = separate(ny_noaa, date, into = c("year", "month", "day"), sep = "-") %>% 
    mutate(prcp = prcp / 10, tmin = as.numeric(tmin) / 10, tmax = as.numeric(tmax) / 10)
```

For snowfall, what are the most commonly observed values? Why?

```{r}
snowfall_daycount = count(ny_noaa, snow, sort = T)
snowfall_daycount[1,1]
```

Make a two-panel plot showing the average temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}
temp_jan_jul = filter(ny_noaa_ymd, month %in% c("01","07")) %>% 
  group_by(id, year, month) %>% 
  summarise(mean_tmax = mean(tmax, na.rm = T)) %>%
  ungroup() %>% 
  mutate(month = recode(month, "01" = "Jan", "07" = "Jul"))

ggplot(temp_jan_jul, aes(x = year, y = mean_tmax,fill = year)) +
  geom_violin(color = "blue", alpha = 0.5) +
  stat_summary(fun.y = median, geom = "point", color = "blue", size = 1) +
  facet_grid(month ~., scales = "free_y") +
  theme(legend.position = "none")
```