---
title: "p8105_hw3_hq2163_hanbo"
author: "Hanbo Qiu"
date: "October 7, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE，
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

#Load the new package. 
library(tidyverse)
library(ggridges)
library(p8105.datasets)
```

##Problem 1

```{r}
#Use code "data(package = "p8105.datasets")" to find the name of target dataset and load it.

data(brfss_smart2010)

#Create an overall_health dataset and focus on the “Overall Health” topic and organize responses as a factor taking levels from “Excellent” to “Poor”.

overall_health = janitor::clean_names(brfss_smart2010) %>% 
  filter(topic == "Overall Health") %>% 
  mutate(response = factor(response, levels = c("Excellent", "Very good", "Good", "Fair", "Poor"), ordered = TRUE))
```

In 2002, which states were observed at 7 locations?
```{r}
dist_loc_02 = filter(overall_health, year == 2002) %>% 
  distinct(locationabbr, locationdesc)

state_loc_7 = count(dist_loc_02, locationabbr) %>% 
  filter(n == 7)
state_loc_7$locationabbr
```
In 2002, the states were observed at 7 locations is `r state_loc_7$locationabbr`.

Make a “spaghetti plot” that shows the number of observations in each state from 2002 to 2010.
```{r}
dist_loc = distinct(overall_health, year, locationabbr, locationdesc)

dist_loc %>% 
  ggplot(aes(x = year, color = locationabbr)) +
  geom_freqpoly(binwidth = 1) +
  scale_x_continuous(breaks = 2002:2010, limits = c(2002,2010))+
  labs(
    title = "The number of observations in each state from 2002 to 2010",
    x = "Year",
    y = "the number of observations in each state",
    caption = "Data from the brfss_smart2010"
  )
```

2002, 2006, and 2010, the mean and standard deviation of the proportion of “Excellent” responses across locations in NY State
```{r}

excellent_ny = filter(overall_health,
                      year %in% c(2002, 2006, 2010), 
                      response == "Excellent", 
                      locationabbr == "NY") %>% 
  group_by(year)

ny_mean_sd = summarise(excellent_ny,
                       excellent_mean = mean(data_value, na.rm = TRUE),
                       excellent_sd = sd(data_value, na.rm = TRUE))
ny_mean_sd
```

each year and state, compute the average proportion and make a five-panel plot
```{r}
response_table = subset(overall_health, select = c(year:locationdesc, response, data_value)) %>% 
  group_by(response, year, locationabbr)

response_mean = summarise(response_table,
                          mean = mean(data_value, na.rm = TRUE))
response_mean %>%  
  ggplot(aes(x = year, y = mean, color = locationabbr)) +
  geom_line() +
  facet_wrap(vars(response), nrow = 5, scales = "free_y", switch = "y")+
  labs(
    title = "Response distribution of state-level averages from 2002-2010",
    x = "Year",
    y = "the average proportion in response category",
    caption = "Data from the brfss_smart2010"
  )
```





